Baseline Drift Estimation for Air Quality Data Using Quantile Trend Filtering

### 2019-03-28 Manuscript/detrendify_aoas.tex
working on week-long application

To terminate our algorithm, we use the stopping criteria described by \cite{boyd2011distributed}. The criteria are based on the primal and dual residuals, which represent the residuals for primal and dual feasibility, respectively. The primal residual at the $m$th iteration,
%	\begin{eqnarray*}
%	r_p^{(q)} & = & \sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert\theta_{j,m}^{(q)} - \M{A}r{\theta}_{j,m}^{(q)}\rVert_2^2},
%	\end{eqnarray*}
\begin{eqnarray*}
	r_{\text{primal}}^{m} & = & \sqrt{\sum_{w=1}^W\lVert\Mn{\Theta}{w}_{m} - \M{\Theta}_m\rVert_{\text{F}}^2},
\end{eqnarray*}
represents the difference between the trend values in the windows and the consensus trend value. The dual residual at the $m$th iteration,
%	\begin{eqnarray*}
%	r_d^{(q)} & = & \gamma\sqrt{\sum_{m=1}^M \sum_{j=1}^J\lVert\M{A}r{\theta}_{j,m}^{(q)} - \M{A}r{\theta}_{j,m}^{(q-1)}\rVert_2^2},
%	\end{eqnarray*}
\begin{eqnarray*}
	r_{\text{dual}}^{m} & = & \gamma\sqrt{\sum_{w=1}^W \lVert\M{\Theta}_m - \M{\Theta}_{m-1}\rVert_{\text{F}}^2},
\end{eqnarray*}
represents the change in the consensus variable from one iterate to the next. The algorithm is stopped when
%	\begin{eqnarray*}
%		r_p^{(q)} & < &\epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\underset{w}{\max}\left[\max
%		\left(\sqrt{\sum_{j=1}^J \lVert\theta_{j,m}^{(q)}\rVert_2^2}, \sqrt{\sum_{j=1}^J \lVert \M{A}r{\theta}_{j,m}^{(q)} \rVert_2^2} \right )\right]\\
%		r_d^{(q)} & < & \epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert \omega_{j,m}^{(q)}\rVert_2^2}
%	\end{eqnarray*}
\begin{eqnarray*}
	r_{\text{primal}}^{m} & < &\epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\,\underset{w}{\max} \left[\max
	\left\{\lVert\Mn{\Theta}{w}_m \rVert_{\text{F}}, \lVert \M{\Theta}_{m} \rVert_{\text{F}} \right\}\right] \\%\lVert \M{\Theta}_{k} \rVert_{\text{F}}} \right )\right]\\
	r_{\text{dual}}^{m} & < & \epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\,\sqrt{\sum_{w=1}^W
		\lVert \Mn{\Omega}{w}_m \rVert_{\text{F}}^2}.
\end{eqnarray*}

@@

We compare the ability of the methods to estimate the true quantiles of $y(t_i)-s(t_i)$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE (\Fig{peaks_rmse}). In this simulation study, our proposed method \texttt{detrend\_eBIC} method substantially outperforms the others. The \texttt{detrend\_cross} method, which is the \texttt{detrend\_eBIC} method fit without the non-crossing constraints, performs similarly for larger quantiles and larger datasets. However, \texttt{detrend\_cross} produces significantly worse estimates for more extreme quantiles ($\tau = 0.01$) and smaller data sets ($\tau = 0.05$ and $n = 500$). These results indicate that even when a single quantile is of interest, simultaneously fitting nearby quantiles and utilizing the non-crossing constraint can improve estimates when data is sparse by using information from nearby quantiles. The \texttt{qsreg} method is comparable to the \texttt{detrend\_eBIC} method on the smaller datasets, but its performance deteriorates as the data size grows. The \texttt{npqw} and \texttt{detrend\_valid} methods both perform poorly on this design.

>>>>

To terminate our algorithm, we use the stopping criteria described by \cite{boyd2011distributed}. The criteria are based on the primal and dual residuals, which represent the residuals for primal and dual feasibility, respectively. The primal residual at the $m$th iteration,

\begin{eqnarray*}
	r_{\text{primal}}^{m} & = & \sqrt{\sum_{w=1}^W\lVert\Mn{\Theta}{w}_{m} - \M{\Theta}_m\rVert_{\text{F}}^2},
\end{eqnarray*}
represents the difference between the trend values in the windows and the consensus trend value. The dual residual at the $m$th iteration,

\begin{eqnarray*}
	r_{\text{dual}}^{m} & = & \gamma\sqrt{\sum_{w=1}^W \lVert\M{\Theta}_m - \M{\Theta}_{m-1}\rVert_{\text{F}}^2},
\end{eqnarray*}
represents the change in the consensus variable from one iterate to the next. The algorithm is stopped when

\begin{eqnarray*}
	r_{\text{primal}}^{m} & < &\epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\,\underset{w}{\max} \left[\max
	\left\{\lVert\Mn{\Theta}{w}_m \rVert_{\text{F}}, \lVert \M{\Theta}_{m} \rVert_{\text{F}} \right\}\right] \\%\lVert \M{\Theta}_{k} \rVert_{\text{F}}} \right )\right]\\
	r_{\text{dual}}^{m} & < & \epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\,\sqrt{\sum_{w=1}^W
		\lVert \Mn{\Omega}{w}_m \rVert_{\text{F}}^2}.
\end{eqnarray*}

@@

We compare the ability of the methods to estimate the true quantiles of $y(t_i)-s(t_i)$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE (\Fig{peaks_rmse}). In this simulation study, our proposed method \texttt{detrend\_eBIC} method substantially outperforms the others. The \texttt{detrend\_Xing} method, which is the \texttt{detrend\_eBIC} method fit without the non-crossing constraints, performs similarly for larger quantiles and larger datasets. However, \texttt{detrend\_Xing} produces significantly worse estimates for more extreme quantiles ($\tau = 0.01$) and smaller data sets ($\tau = 0.05$ and $n = 500$). These results indicate that even when a single quantile is of interest, simultaneously fitting nearby quantiles and utilizing the non-crossing constraint can improve estimates when data is sparse by using information from nearby quantiles. The \texttt{qsreg} method is comparable to the \texttt{detrend\_eBIC} method on the smaller datasets, but its performance deteriorates as the data size grows. The \texttt{npqw} and \texttt{detrend\_valid} methods both perform poorly on this design.
